### 2020-08-25 A03_slope_impl_details.tex

% appendix 
% considerations for cyclic boundary conditions
Sloping surface requires cyclic boundary conditions. There are two possible approaches in the sloping surface case:
(a) Slope offset: Choose initial conditions in which isopycnals follow isobars. When water is cycled across boundaries, apply a temperature offset determined by the slope and the initial temperature gradient. Apply a salinity offset as well (not implemented in PALM v0).
(b) No offset: Choose initial conditions in which isopycnals are parallel to the surface. Do not apply any temperature or salinity offset to water cycled across boundaries.