### 2020-08-25 custom_header.tex

\documentclass[letterpaper,10pt]{report}

\usepackage{ifxetex,ifluatex} % provides way to check if document is being processed with xetex or luatex

\usepackage[colorlinks = true,
unicode=true,
linkcolor = Black,
urlcolor  = Black,
citecolor = Black,
anchorcolor = blue]{hyperref}

\usepackage[english]{babel} % language support

% --- page layout ---

\usepackage{lastpage} % allows user to refer to last page with label LastPage
\usepackage{titlesec} % options for title, headers, contents

%header
\usepackage{fancyhdr} % customize headers and footers
\pagestyle{fancy}
\fancyhf{}
\rhead{Carolyn Begeman, PALM development, Page \thepage \thinspace \thinspace of \pageref{LastPage}}
\lhead{}
\renewcommand{\headrulewidth}{0pt}

% page dimensions
\usepackage{geometry}
\geometry{letterpaper, top=0.75in, left=1in, total={6.5in,9.5in}, headheight=12pt, headsep=.1in} % topmargin=-23pt,

\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother

% set paragraph settings
\IfFileExists{parskip.sty}{%
	\usepackage{parskip}
}{% else
	\setlength{\parindent}{0pt}
	\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightitemize}{%
	\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\setcounter{secnumdepth}{3} % set maximum depth of subsections

%\usepackage{caption}
%\captionsetup[table]{skip=8pt]}

% --- font-related packages ---

\usepackage{mathptmx} % make Times default font, with math font
\usepackage{lmodern} % load Latin Modern family of fonts
\usepackage{amssymb,amsmath} % define AMS symbol fonts
\usepackage[usenames,dvipsnames]{color} % color model

% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

% use microtype if available
\IfFileExists{microtype.sty}{%
	\usepackage[]{microtype}
	\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\hypersetup{
	pdfborder={0 0 0},
	breaklinks=true}
%\urlstyle{same}  % don't use monospace font for urls
%\usepackage[utf8]{inputenc} % specify input encoding, \DeclareUnicodeCharacter

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\else % if luatex or xelatex
\ifxetex
\usepackage{mathspec}
\else
\usepackage{fontspec}
\fi
\defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi

% set fontsize for section and subsection
\newcommand{\secfnt}{\fontsize{12}{14}}
\titlespacing\section{0pt}{0pt plus 2pt minus 2pt}{0pt plus 2pt minus 2pt}
\newcommand{\ssecfnt}{\fontsize{12}{14}}
\titlespacing\subsection{0pt}{0pt plus 2pt minus 2pt}{0pt plus 2pt minus 2pt}
\renewcommand{\baselinestretch}{1}

% settings for text that designates code
\definecolor{light-gray}{gray}{0.95} 
\newcommand{\code}[1]{\texttt{#1}} %{\colorbox{light-gray}

% --- itemizes ---
\usepackage{enumitem,amssymb}
%\newitemize{todoitemize}{itemize}{2}
%\setitemize[todoitemize]{label=$\square$}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
	\hspace{-2.5pt}}
\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}

% --- graphics ---

\usepackage{graphicx,grffile}
\usepackage{makecell}

% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

%\graphicspath{ {} }

%\date{} %replace date with text \date{text}