### 2021-01-18 A_suppl_fig.tex

\begin{figure}
    \centering
    \includegraphics{}
    \caption{Timeseries of maximum vertical velocity fluctuation per hour for thermal driving cases}
    \label{fig:max_w2_timeseries}
\end{figure}