### 2021-04-19 A_suppl_fig.tex
Updates from Overleaf

\begin{figure}
    \centering
    \begin{minipage}{0.5\textwidth}
        \includegraphics{Figures/w2_max_cmp_dT_t.png}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
        \includegraphics{Figures/w2_max_cmp_dalpha_t.png}
    \end{minipage}
    \caption{Timeseries of maximum vertical velocity fluctuation per hour for (a) thermal driving cases and (b) slope cases.}
    \label{fig:max_w2_timeseries}
\end{figure}

\begin{figure}
    \centering
    \includegraphics{Figures/e_res_cmp_dpdy_dpdx_t.png}
    \caption{Resolved turbulent kinetic energy as a function of pressure gradient orientation under 0.1$^{\circ}$ slope.}
    \label{fig:e_orientation}
\end{figure}

\begin{figure}
    \centering
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 10cm 0 0},clip, width=\textwidth]{Figures/heatflux_cmp_dT_40hr_tav1_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 10cm 0 0},clip, width=\textwidth]{Figures/heatflux_cmp_dT_48hr_tav1_z_profile.png}
    \end{minipage}
    \caption{Thermal driving simulations: heat flux}
    \label{fig:Heatflux_profiles}
\end{figure}